Prevailing scenarios of functional change in anthropocene bird and mammal communities

### 2023-06-22 paper/GEB Submission/GEB_title_page.tex

---
title: Prevailing scenarios of functional change in anthropocene bird and mammal communities
authors:
  - name: "Kari E. A. Norman"
    affiliation: Université de Montréal
    location: Montréal, Quebec
    email: kari.norman@berkeley.edu
  - name: "Perry de Valpine"
    affiliation: University of California, Berkeley
    location: Berkeley, California
    email: pdevalpine@berkeley.edu
  - name: "Carl Boettiger"
    affiliation: University of California, Berkeley
    location: Berkeley, California
    email: cboettig@berkeley.edu
abstract: |
  Aim: Despite unprecedented environmental change due to anthropogenic pressure, recent work has found increasing species turnover but no overall trend in species diversity through time at the local scale. Functional diversity provides a potentially powerful alternative approach for understanding community composition by linking shifts in species identity to mechanisms of ecosystem processes. Here we present the first multi-taxa, multi-system analysis of functional change through time.

@@

  Main Conclusions: General trends indicate that on the aggregate one type of functional shift is not more prevelant than the other across many taxa, biomes, and realms. At the study-level, there were also a substantial number of time series exhibiting no species or functional change, however the majority of studies showed a shift in a species or functional metric. With no one prevailing scenario of change, it will be critical to link change scenarios to drivers of change, particularly to identify communities with capacity to resist drivers from those not experiencing substantial pressure from a driver.
journal: "Ecology Letters"
date: "`r Sys.Date()`"
bibliography: refs.bib
csl: pnas.csl
layout: 3p
header-includes:
   - \usepackage{lineno}
   - \linenumbers
output: 
  rticles::arxiv_article:
    keep_tex: true
#  word_document: default
editor_options: 
  markdown: 
    wrap: sentence
---

# Data Availability

Code for the analyses in this chapter is archived on Zenodo at [10.5281/zenodo.5514334](https://doi.org/10.5281/zenodo.5514334).

Data products are also archived on Zenodo at [10.5281/zenodo.6499442](https://doi.org/10.5281/zenodo.6499442).

@@

# Funding

Support for this work was provided by U.S. Department of Energy through the Computational Sciences Graduate Fellowship (DE-FG02-97ER25308) awarded to KEAN. CB acknowledges support from NSF CAREER Award #1942280 & NIFA project CA-B-INS-0162-H.

# Biosketch

Kari Norman’s research is generally focused on understanding the impacts of anthropogenic change on biodiversity. She brings together existing data sources to understand broadscale change and develops software tools to facilitate robust synthesis workflows.

>>>>

\documentclass{article}

\usepackage{arxiv}

\usepackage[utf8]{inputenc} % allow utf-8 input
\usepackage[T1]{fontenc}    % use 8-bit T1 fonts
\usepackage{lmodern}        % https://github.com/rstudio/rticles/issues/343
\usepackage{hyperref}       % hyperlinks
\usepackage{url}            % simple URL typesetting
\usepackage{booktabs}       % professional-quality tables
\usepackage{amsfonts}       % blackboard math symbols
\usepackage{nicefrac}       % compact symbols for 1/2, etc.
\usepackage{microtype}      % microtypography
\usepackage{graphicx}

\title{Prevailing scenarios of functional change in anthropocene bird
and mammal communities}

\author{
    Kari E. A. Norman
   \\
    Université de Montréal \\
  Montréal, Quebec \\
  \texttt{\href{mailto:kari.norman@berkeley.edu}{\nolinkurl{kari.norman@berkeley.edu}}} \\
   \And
    Perry de Valpine
   \\
    University of California, Berkeley \\
  Berkeley, California \\
  \texttt{\href{mailto:pdevalpine@berkeley.edu}{\nolinkurl{pdevalpine@berkeley.edu}}} \\
   \And
    Carl Boettiger
   \\
    University of California, Berkeley \\
  Berkeley, California \\
  \texttt{\href{mailto:cboettig@berkeley.edu}{\nolinkurl{cboettig@berkeley.edu}}} \\
  }

% tightlist command for lists without linebreak
\providecommand{\tightlist}{%
  \setlength{\itemsep}{0pt}\setlength{\parskip}{0pt}}

\usepackage{lineno}
\linenumbers
\begin{document}
\maketitle

\begin{abstract}
Aim: Despite unprecedented environmental change due to anthropogenic
pressure, recent work has found increasing species turnover but no
overall trend in species diversity through time at the local scale.
Functional diversity provides a potentially powerful alternative
approach for understanding community composition by linking shifts in
species identity to mechanisms of ecosystem processes. Here we present
the first multi-taxa, multi-system analysis of functional change through
time.

@@

Main Conclusions: General trends indicate that on the aggregate one type
of functional shift is not more prevelant than the other across many
taxa, biomes, and realms. At the study-level, there were also a
substantial number of time series exhibiting no species or functional
change, however the majority of studies showed a shift in a species or
functional metric. With no one prevailing scenario of change, it will be
critical to link change scenarios to drivers of change, particularly to
identify communities with capacity to resist drivers from those not
experiencing substantial pressure from a driver.
\end{abstract}

\hypertarget{data-availability}{%
\section{Data Availability}\label{data-availability}}

Code for the analyses in this chapter is archived on Zenodo at
\href{https://doi.org/10.5281/zenodo.5514334}{10.5281/zenodo.5514334}.

Data products are also archived on Zenodo at
\href{https://doi.org/10.5281/zenodo.6499442}{10.5281/zenodo.6499442}.

@@

\hypertarget{funding}{%
\section{Funding}\label{funding}}

Support for this work was provided by U.S. Department of Energy through
the Computational Sciences Graduate Fellowship (DE-FG02-97ER25308)
awarded to KEAN. CB acknowledges support from NSF CAREER Award \#1942280
\& NIFA project CA-B-INS-0162-H.

\bibliographystyle{unsrt}
\bibliography{refs.bib}

\end{document}
